---
title: "RopenMeteo"
output: github_document
---

<!-- badges: start -->
  [![R-CMD-check](https://github.com/FLARE-forecast/RopenMeteo/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/FLARE-forecast/RopenMeteo/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

R wrappers for APIs on Open-Meteo project.  The Open-Meteo is a amazing project that streamlines the access to a range of publicly historical and forecasted meteorology data from agencies across the world.  The free access tier allows for 10,000 API calls per day.  The paid tiers increase the number of daily API calls (support for paid APIs in this package is pending). Learn more about the Open-Meteo project at their website ([https://open-meteo.com]) and consider supporting their efforts.

Open-Meteo citation: Zippenfenig, Patrick. (2023). Open-Meteo.com Weather API (0.2.69). Zenodo. https://doi.org/10.5281/zenodo.8112599

The package includes additional functionally to facilitate the use in mechanistic environmental/ecological models.  This includes the calculation of longwave radiation (not provided through the API) from air temperature and cloud cover, the writing of output to the format required by the General Lake Model (GLM), and the conversion to the standard used in the NEON Ecological Forecasting Challenge that is run by the Ecological Initiative Research Coordination Network ([https://neon4cast.org]).  Future functionally includes the temporal downscaling of the daily climate projection output and the 6-hourly seasonal forecast to the hourly time step.

The package uses a long format standard with the following columns

```
datetime = date and time of forecasted value
reference_datetime = the date and time of the beginning of the forecast (horizon = 0). Does not apply to historical weather.
model_id = id of model that generated the forecast
ensemble = ensemble member number (only for ensemble weather and seasonal forecasts)
variable = forecasted variable
prediction = forecasted value
unit = units of the variable
```

## Install

```{r message = FALSE, warning = FALSE}
remotes::install_github("FLARE-forecast/RopenMeteo")
```

```{r message = FALSE}
library(tidyverse)
```

## Weather forecasts

The open-meteo project combines the the best models for each location across the globe to provide the best possible forecast.  open-meteo defines this as `model = "generic"`.

[https://open-meteo.com/en/docs]

```{r}
df <- RopenMeteo::get_forecast(latitude = 37.30,
                               longitude = -79.83,
                               forecast_days = 7, 
                               past_days = 2, 
                               model = "generic",
                               variables = c("temperature_2m"))
head(df)
```

```{r}
df |> 
  mutate(variable = paste(variable, unit)) |> 
  ggplot(aes(x = datetime, y = prediction)) + 
  geom_line(color = "#F8766D") + 
  geom_vline(aes(xintercept = reference_datetime)) + 
  facet_wrap(~variable, scale = "free")
```

## Ensemble Weather Forecasts

Ensemble forecasts from individual models are available. 

[https://open-meteo.com/en/docs/ensemble-api]

```{r}
df <- RopenMeteo::get_ensemble_forecast(
  latitude = 37.30,
  longitude = -79.83,
  forecast_days = 7,
  past_days = 2,
  model = "gfs_seamless",
  variables = c("temperature_2m"))
head(df)
```

```{r}
df |> 
  mutate(variable = paste(variable, unit)) |> 
  ggplot(aes(x = datetime, y = prediction, color = ensemble)) + 
  geom_line() + 
  geom_vline(aes(xintercept = reference_datetime)) + 
  facet_wrap(~variable, scale = "free", ncol = 2)
```

Options for models and variables are at https://open-meteo.com/en/docs/ensemble-api

Note that `ecmwf_ifs04` does not include solar radiation.  

List of global model ids: 

```
icon_seamless, icon_global, gfs_seamless, gfs025, gfs05, ecmwf_ifs04, gem_global
```

### Use with the General Lake Model

We have included functions that allow the output to be used with the General Lake Model ([https://doi.org/10.5194/gmd-12-473-2019]).
Since the open-meteo models do not include longwave radiation, the package provides a function to calculate it from the cloud cover and air temperature.

GLM requires a set of variables that are provided

```{r}
df <- RopenMeteo::get_ensemble_forecast(
  latitude = 37.30,
  longitude = -79.83,
  forecast_days = 7,
  past_days = 2,
  model = "gfs_seamless",
  variables = c(
    "relativehumidity_2m",
    "precipitation",
    "windspeed_10m",
    "cloudcover",
    "temperature_2m",
    "shortwave_radiation"))
head(df)
```

```{r}
df |> 
  mutate(variable = paste(variable, unit)) |> 
  ggplot(aes(x = datetime, y = prediction, color = ensemble)) + 
  geom_line() + 
  geom_vline(aes(xintercept = reference_datetime)) + 
  facet_wrap(~variable, scale = "free", ncol = 2)
```

The following converts to GLM format

```{r}
path <- tempdir()
df |> 
    RopenMeteo::add_longwave() |>
    RopenMeteo::write_glm_format(path = path)
  head(read.csv(list.files(path = path, full.names = TRUE, pattern = ".csv")[1]))
```

### Converting to Ecological Forecasting Initative convention

The standard used in the NEON Ecological Forecasting Challenge is slightly different from the standard in this package.  It uses the column `parameter` for ensemble because the Challenge standard allows the flexibility to use parametric distributions (i.e., normal distribution `mean` and `sd`) in the same standard as a ensemble (or sample) forecast.  The `family` column defines the distribution (here `family` = `ensemble`).

The EFI standard also follows CF-conventions so the variable names are converted to be CF compliant. 

The output from `RopenMeteo::convert_to_efi_standard()` is the same as the output from `neon4cast::stage2()`

Learn more about `neon4cast::stage2()` here: [https://projects.ecoforecast.org/neon4cast-docs/Shared-Forecast-Drivers.html]

```{r}
df |>
  RopenMeteo::add_longwave() |>
  RopenMeteo::convert_to_efi_standard()
```

Note that `neon4cast::stage3()` is similar to

```{r}
df |>
  RopenMeteo::add_longwave() |>
  RopenMeteo::convert_to_efi_standard() |> 
  filter(datetime < reference_datetime)
```

With the only difference that the number of days is equal to the `past_days` in the call to `RopenMeteo::get_ensemble_forecast()`.  The max `past_days` from open-meteo is ~60 days.

## Historical Weather

If you need more historical days for model calibration and testing, historical data are available through open-meteo's historical weather API.

[https://open-meteo.com/en/docs/historical-weather-api] 

```{r}
df <- RopenMeteo::get_historical_weather(
  latitude = 37.30,
  longitude = -79.83,
  start_date = "2023-01-01",
  end_date = Sys.Date(),
  variables = c("temperature_2m")) 
tail(df |> na.omit())
```

Notice the delay of ~7 days. 

```{r}
df |> 
  mutate(variable = paste(variable, unit)) |> 
  ggplot(aes(x = datetime, y = prediction)) + 
  geom_line(color = "#F8766D") + 
  geom_vline(aes(xintercept = lubridate::with_tz(Sys.time(), tzone = "UTC"))) + 
  facet_wrap(~variable, scale = "free")
```

## Seasonal Forecasts

Weather forecasts for up to 9 months in the future are available from the NOAA Climate Forecasting System

[https://open-meteo.com/en/docs/seasonal-forecast-api]

```{r}
df <- RopenMeteo::get_seasonal_forecast(
  latitude = 37.30,
  longitude = -79.83,
  forecast_days = 274,
  past_days = 5,
  variables = c("temperature_2m"))
head(df)
```

```{r}
df |> 
  mutate(variable = paste(variable, unit)) |> 
  ggplot(aes(x = datetime, y = prediction, color = ensemble)) + 
  geom_line() + 
  geom_vline(aes(xintercept = reference_datetime)) +
  facet_wrap(~variable, scale = "free")
```

## Climate Projections

Climate projections from different models are available through 2050.

[https://open-meteo.com/en/docs/climate-api]

```{r}
df <- RopenMeteo::get_climate_projections(
  latitude = 37.30,
  longitude = -79.83,
  start_date = Sys.Date(),
  end_date = Sys.Date() + lubridate::years(1),
  model = "EC_Earth3P_HR",
  variables = c("temperature_2m_mean"))
head(df)
```

### Multiple climate models

```{r}
models <- c("CMCC_CM2_VHR4","FGOALS_f3_H","HiRAM_SIT_HR","MRI_AGCM3_2_S","EC_Earth3P_HR","MPI_ESM1_2_XR","NICAM16_8S")

df <- purrr::map_df(models, function(model){
  RopenMeteo::get_climate_projections(
    latitude = 37.30,
    longitude = -79.83,
    start_date = Sys.Date(),
    end_date = Sys.Date() + lubridate::years(1),
    model = model,
    variables = c("temperature_2m_mean"))
  })
  
```

```{r}
df |> 
    mutate(variable = paste(variable, unit)) |> 
    ggplot(aes(x = datetime, y = prediction, color = model_id)) + 
    geom_line() +
    facet_wrap(~variable, scale = "free")
```


