---
title: "Example Usage"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example Usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r message = FALSE}
library(ggplot2)
library(dplyr)
library(purrr)
library(tibble)
library(ropenmeteo)
library(lubridate)
library(readr)
```

## Weather forecasts

The open-meteo project combines the best models for each location across the
globe to provide the best possible forecast. open-meteo refers to this as
`model = "generic"`. The maximum forecast horizon is 16 days.

<https://open-meteo.com/en/docs>

```{r}
df <- get_forecast(latitude = 37.30,
                   longitude = -79.83,
                   forecast_days = 7,
                   past_days = 2,
                   model = "generic",
                   variables = c("temperature_2m"))
head(df)
```

```{r}
df |>
  mutate(variable = paste(variable, unit)) |>
  ggplot(aes(x = datetime, y = prediction)) +
  geom_line(color = "#F8766D") +
  geom_vline(aes(xintercept = reference_datetime)) +
  facet_wrap(~variable, scale = "free")
```

### Additional forecast models

Besides `"generic"`, model-specific endpoints are available. These can be
useful when you need a particular agency's output or regional high-resolution
data.

| Model ID        | Agency                        | Coverage          |
|-----------------|-------------------------------|-------------------|
| `"gfs"`         | NOAA (US)                     | Global            |
| `"ecmwf"`       | ECMWF (EU)                    | Global, 15 days   |
| `"meteofrance"` | Météo-France                  | Global/Europe     |
| `"dwd"`         | DWD (Germany, ICON)           | Global            |
| `"gem"`         | Environment Canada            | Global, 10 days   |
| `"jma"`         | JMA (Japan)                   | Global, 11 days   |
| `"metno"`       | MET Norway                    | Scandinavia only  |
| `"kma"`         | KMA (Korea)                   | Global, 12 days   |
| `"bom"`         | BOM ACCESS-G (Australia)      | Global, 7 days    |
| `"ukmo"`        | UK Met Office                 | Global, 7 days    |

```{r eval = FALSE}
df_ecmwf <- get_forecast(latitude = 37.30,
                         longitude = -79.83,
                         forecast_days = 7,
                         past_days = 2,
                         model = "ecmwf",
                         variables = c("temperature_2m"))
```

## Ensemble Weather Forecasts

Ensemble forecasts provide probabilistic output from multiple model runs and
are available for up to 35 days ahead.

<https://open-meteo.com/en/docs/ensemble-api>

```{r eval = FALSE}
df <- get_ensemble_forecast(
  latitude = 37.30,
  longitude = -79.83,
  forecast_days = 7,
  past_days = 2,
  model = "gfs_seamless",
  variables = c("temperature_2m"))
head(df)
```

```{r forecast-plot, eval=FALSE}
df |>
  mutate(variable = paste(variable, unit)) |>
  ggplot(aes(x = datetime, y = prediction, color = ensemble)) +
  geom_line() +
  geom_vline(aes(xintercept = reference_datetime)) +
  facet_wrap(~variable, scale = "free", ncol = 2)
```

Available ensemble models include:

```
icon_seamless_eps, icon_global_eps, icon_eu_eps, icon_d2_eps,
gfs_seamless_eps, gfs025_eps, gfs05_eps,
ecmwf_ifs025, ecmwf_aifs025,
gem_global, bom_access_global,
ukmo_global_20km, ukmo_uk_2km
```

Note: `ecmwf_ifs025` does not include shortwave radiation.

### Use with the General Lake Model

We have included functions that allow the output to be used with the General
Lake Model (<https://doi.org/10.5194/gmd-12-473-2019>). Since the open-meteo
models do not include longwave radiation, the package provides a function to
calculate it from cloud cover and air temperature.

`glm_variables()` returns the variable set required by GLM for each product
and time step:

```{r}
df <- get_ensemble_forecast(
  latitude = 37.30,
  longitude = -79.83,
  forecast_days = 7,
  past_days = 2,
  model = "gfs_seamless",
  variables = glm_variables(product = "ensemble_forecast",
                            time_step = "hourly"))
head(df)
```

```{r ensemble-plot}
df |>
  mutate(variable = paste(variable, unit)) |>
  ggplot(aes(x = datetime, y = prediction, color = ensemble)) +
  geom_line() +
  geom_vline(aes(xintercept = reference_datetime)) +
  facet_wrap(~variable, scale = "free", ncol = 2)
```

The following converts to GLM format:

```{r}
path <- tempdir()
df |>
  add_longwave() |>
  write_glm_format(path = path)
head(read_csv(list.files(path = path, full.names = TRUE, pattern = ".csv")[1],
              show_col_types = FALSE))
```

### Converting to Ecological Forecasting Initiative convention

The standard used in the NEON Ecological Forecasting Challenge differs
slightly from this package's standard. It uses `parameter` instead of
`ensemble` because the Challenge standard accommodates both ensemble
(sample) and parametric distribution forecasts in the same format. The
`family` column defines the distribution type (here `family = "ensemble"`).

The EFI standard also follows CF conventions, so variable names are converted
to be CF-compliant.

The output from `convert_to_efi_standard()` matches the format of
`neon4cast::stage2()`. Learn more at
<https://projects.ecoforecast.org/neon4cast-docs/Shared-Forecast-Drivers.html>.

```{r}
df |>
  add_longwave() |>
  convert_to_efi_standard()
```

Note that `neon4cast::stage3()` is equivalent to filtering to the observed
period:

```{r}
df |>
  add_longwave() |>
  convert_to_efi_standard() |>
  filter(datetime < reference_datetime)
```

The number of days in the observed period equals `past_days` in the call to
`get_ensemble_forecast()`. The maximum `past_days` from open-meteo is 92 days.

## Historical Weather

If you need more historical days for model calibration and testing, historical
data are available through open-meteo's ERA5 reanalysis API. ERA5 is available
from 1940-01-01 to approximately 5 days before the current date.

<https://open-meteo.com/en/docs/historical-weather-api>

```{r}
df <- get_historical_weather(
  latitude = 37.30,
  longitude = -79.83,
  start_date = "2023-01-01",
  end_date = Sys.Date() - lubridate::days(5),
  variables = c("temperature_2m"))
tail(df |> na.omit())
```

```{r hist-plot}
df |>
  mutate(variable = paste(variable, unit)) |>
  ggplot(aes(x = datetime, y = prediction)) +
  geom_line(color = "#F8766D") +
  geom_vline(aes(xintercept = lubridate::with_tz(Sys.time(), tzone = "UTC"))) +
  facet_wrap(~variable, scale = "free")
```

## Seasonal Forecasts

Seasonal forecasts covering up to 7 months ahead are provided by ECMWF.
The default `"ecmwf_seasonal_seamless"` model blends EC46 (updated daily,
46-day horizon) with SEAS5 (updated monthly, 7-month horizon) for the
best available coverage at any lead time.

<https://open-meteo.com/en/docs/seasonal-forecast-api>

```{r}
df <- get_seasonal_forecast(
  latitude = 37.30,
  longitude = -79.83,
  forecast_days = 30,
  past_days = 5,
  variables = c("temperature_2m"))
head(df)
```

```{r seasonal-plot}
df |>
  mutate(variable = paste(variable, unit)) |>
  ggplot(aes(x = datetime, y = prediction, color = ensemble)) +
  geom_line() +
  geom_vline(aes(xintercept = reference_datetime)) +
  facet_wrap(~variable, scale = "free")
```

### Downscaling from 6-hourly to hourly time step

The seasonal API delivers data at 6-hourly intervals. The package provides
`six_hourly_to_hourly()` to downscale to hourly using linear interpolation
for most variables and solar geometry for shortwave radiation.

```{r}
df <- get_seasonal_forecast(
  latitude = 37.30,
  longitude = -79.83,
  forecast_days = 30,
  past_days = 5,
  variables = glm_variables(product = "seasonal_forecast",
                            time_step = "6hourly"))
```

```{r downscale-plot}
df |>
  six_hourly_to_hourly(latitude = 37.30, longitude = -79.83,
                       use_solar_geom = TRUE) |>
  mutate(variable = paste(variable, unit)) |>
  ggplot(aes(x = datetime, y = prediction, color = ensemble)) +
  geom_line() +
  geom_vline(aes(xintercept = reference_datetime)) +
  facet_wrap(~variable, scale = "free", ncol = 2)
```

## Climate Projections

Daily climate projections from seven high-resolution CMIP6 models are
available from 1950 through 2050.

Note that shortwave radiation units differ from the forecast APIs: the climate
API returns `shortwave_radiation_sum` in MJ/m² (daily total) rather than
instantaneous W/m².

<https://open-meteo.com/en/docs/climate-api>

```{r}
df <- get_climate_projections(
  latitude = 37.30,
  longitude = -79.83,
  start_date = Sys.Date(),
  end_date = Sys.Date() + lubridate::years(1),
  model = "EC_Earth3P_HR",
  variables = c("temperature_2m_mean"))
head(df)
```

```{r climate-plot}
df |>
  mutate(variable = paste(variable, unit)) |>
  ggplot(aes(x = datetime, y = prediction)) +
  geom_line(color = "#F8766D") +
  facet_wrap(~variable, scale = "free")
```

## Downloading multiple sites or models

### Multiple models

```{r}
models <- c("CMCC_CM2_VHR4", "FGOALS_f3_H", "HiRAM_SIT_HR",
            "MRI_AGCM3_2_S", "EC_Earth3P_HR", "MPI_ESM1_2_XR", "NICAM16_8S")

df <- map_df(models, function(model) {
  get_climate_projections(
    latitude = 37.30,
    longitude = -79.83,
    start_date = Sys.Date(),
    end_date = Sys.Date() + lubridate::years(1),
    model = model,
    variables = c("temperature_2m_mean"))
})
```

```{r multi-climate-plot}
df |>
  mutate(variable = paste(variable, unit)) |>
  ggplot(aes(x = datetime, y = prediction, color = model_id)) +
  geom_line() +
  facet_wrap(~variable, scale = "free")
```

### Multiple sites

The optional `site_id` argument labels rows for each location, making it easy
to combine results from multiple sites.

```{r}
sites <- tibble(site_id = c("fcre", "sunp"),
                latitude = c(37.30, 43.39),
                longitude = c(-79.83, -72.05))

df <- map_df(1:nrow(sites), function(i, sites) {
  get_climate_projections(
    latitude = sites$latitude[i],
    longitude = sites$longitude[i],
    site_id = sites$site_id[i],
    start_date = Sys.Date(),
    end_date = Sys.Date() + lubridate::years(1),
    model = "MPI_ESM1_2_XR",
    variables = c("temperature_2m_mean"))
}, sites)
head(df)
```

```{r multi-site-plot}
df |>
  mutate(variable = paste(variable, unit)) |>
  ggplot(aes(x = datetime, y = prediction, color = site_id)) +
  geom_line() +
  facet_wrap(~variable, scale = "free")
```

### Converting from daily to hourly time step

Photosynthesis responds non-linearly to shortwave radiation, so applying a
daily mean to every hour underestimates peak midday production. `daily_to_hourly()`
uses solar geometry to redistribute the daily radiation total across daylight
hours, and divides precipitation sums evenly across 24 hours. All other
variables have their daily mean applied to each hour.

```{r}
df <- get_climate_projections(
  latitude = 37.30,
  longitude = -79.83,
  start_date = Sys.Date(),
  end_date = Sys.Date() + lubridate::years(1),
  model = "EC_Earth3P_HR",
  variables = glm_variables(product = "climate_projection", time_step = "daily"))
```

```{r daily-to-hourly-plot}
df |>
  daily_to_hourly(latitude = 37.30, longitude = -79.83) |>
  mutate(variable = paste(variable, unit)) |>
  ggplot(aes(x = datetime, y = prediction)) +
  geom_line(color = "#F8766D") +
  facet_wrap(~variable, scale = "free", ncol = 2)
```
